Proof nets for first-order additive linear logic

### 2018-01-20 ALL1.tex
prose up to page 10

\subjclass{F.4.1 [Mathematical logic and formal languages]: Mathematical logic--Proof theory}
\keywords{Linear logic, First-order logic, Proof nets}

@@

Witness nets are canonical for first-order additive sequent calculus. Unification nets in addition factor out any inessential choice for existential witnesses. 
Both notions of proof net are defined through coalescence (the additive version of multiplicative contractibility), and for witness nets a geometric correctness criterion is provided. Both feature a global, one-shot composition operation.
\end{abstract}

@@

Additive linear logic $(\all)$ is the logic of product $(\*)$ and sum $(\+)$. Proofs, over a sequent $A\vdash B$ or, here, $\seq{\dual A}B$, are built from the canonical morphisms: projections, injections, and diagonals. These are captured naturally in \emph{proof nets} \cite{Girard-1987,Hughes-vanGlabbeek-2005}, which consist of a set of \emph{links} between subformulas in $\dual A$ and in $B$, satisfying a simple geometric correctness condition.
\[
\begin{array}{c@{\qquad}c@{\qquad}c@{\qquad}c}
	\begin{tikzpicture}[net]
		\formula[y=1.2]{{\dual a}}
		\formula[y=0]{a\*a}
		\Vlink[red]{[-2]1,1}
		\Vlink[red]{[2]1,3}
	\end{tikzpicture}
&
	\begin{tikzpicture}[net]
		\formula[y=1.2]{{\dual a}}
		\formula[y=0]{a\+b}
		\Vlink[red]{1,1}
	\end{tikzpicture}
&
	\begin{tikzpicture}[net]
		\formula[y=1.2]{{\dual a}\+{\dual b}}
		\formula[y=0]{b\*a}
		\Vlink[red]{1,3}
		\Vlink[red]{3,1}
	\end{tikzpicture}
&
	\begin{tikzpicture}[net]
		\formula[y=1.2]{*{\phantom(}{\dual a}*{\phantom)}
		 {{\+}}*({\dual b}*{\phantom)}
		 {{\+}}*{\phantom(}{\dual c}*)}
		\formula[y=0]{*(a*{\phantom)}
		{{\*}}*{\phantom(}b*)
		{{\*}}*{\phantom(}c*{\phantom)}}
		\Vlink[red]{1,1}
		\Vlink[red]{3,3}
		\Vlink[red]{5,5}
	\end{tikzpicture}
\\
	\text{\small diagonal}
&	\text{\small injection}
&	\text{\small symmetry}
&	\text{\small associativity}
\end{array}
\]

The logic is combinatorially rich, yet well-behaved and tractable: proof search \cite{Galmiche-Marion-1995,Galmiche-2000} and proof net correctness \cite{Heijltjes-Hughes-2015} are linear in $|A|\times|B|$ (with $|A|$ the size of the syntax tree of $A$); proof nets remain canonical and equally tractable when extended with the two units \cite{Heijltjes-2011,Heijltjes-Hughes-2015}; and the first-order case is merely NP-complete \cite{Heijltjes-Hughes-2015}.

At the same time, the ubiquity of its main operations, product and sum, mean the ideas and results it garners are widely applicable. It serves as a microcosmos for important ideas and observations: it fostered the \emph{connections method} of proof search \cite{Galmiche-2000}; it demonstrated the \emph{Blass problem} \cite{Abramsky-2003} of game semantics, that sequential strategies do not in general have associative composition; and it is at the root of the study of fixed points in linear logic \cite{Santocanale-2002a}. The logic describes two-way communication~\cite{Cockett-Santocanale-2009}, and as such is a core part of \emph{session types}~\cite{Honda-1993}.

{\all} is of course part of {\mall} (multiplicative-additive linear logic), and its lessons are clearly visible in the second author's canonical proof nets \cite{Hughes-vanGlabbeek-2005} for that fragment, as well as the first and second author's locally canonical \emph{conflict nets}~\cite{Hughes-Heijltjes-2016}. But its proof nets also appear as the \emph{skew fibrations} in the second author's \emph{combinatorial proofs} for classical logic~\cite{Hughes-2006-proofs}, and in the third authors study of the \emph{medial rule} for classical logic~\cite{str:RTA07}. To prepare the ground for the extension of \emph{combinatorial proofs} to first-order classical logic is one motivation for the current presentation.

The main content of first-order proof, and the central challenge for first-order additive proof nets, is the \emph{witness assignment} to existential quantifiers. There are two, related obstacles in particular. First, the standard approach to evaluate a formula $\exists x.A$ is by a substitution $A[t/x]$, which destroys the subformula $A$.%
%
\footnote{Given $A[t/x]$ and $t$, it is not possible to unambiguously recover $A$, which may contain other occurrences of $t$ not originating in the substitution $[t/x]$.}
%
But the current, successful formulation of proof nets, as a set of links over a two-formula sequent, requires subformulas to remain intact.

Second, the same existential quantification $\exists x.A$ may occur in different \emph{slices} (branches of a proof above a product), with different witnesses, as in the below proof. But the formulas--plus--links design has exactly one instance of $\exists x.A$.
\[
\vc{
	\infer[\Qrr*]{\Seq{\exists x.A}{B\*C}}{
	 \infer[\Qrr?s]{\Seq{\exists x.A}B}{\Seq{A[s/x]}B}
	 &
	 \infer[\Qrr?t]{\Seq{\exists x.A}C}{\Seq{A[t/x]}C}
   }}
~.
\]

We advance two solutions, both simple and radical. \defn{Witness nets} replace the \emph{implicit} substitutions of sequent proofs with \emph{explicit} substitutions, recorded at each link. An example is given below left. Second, \defn{unification nets} omit existential witnesses altogether, and reconstruct them by unification. The example is below right.
\[
\begin{tikzpicture}[net]
	\formula[y=1.2]{\forall*{x.}\exists*{y.~}{\dual P}*{(y)}\+{\dual Q}*{(x,y)}}
	\formula[y=0]{\exists*{z.~}P*{(f(z))}\*Q*{(z,t)}}
	\Vlink[red,l,label={$\scriptstyle[f(x)/y\,,\,x/z]$~~}]{3,2}
	\Vlink[blue,r,label={~$\scriptstyle[t/y\,,\,x/z]$}]{5,4}
\end{tikzpicture}
\qquad\qquad
\begin{tikzpicture}[net]
	\formula[y=1.2]{\forall*{x.}\exists*{y.~}{\dual P}*{(y)}\+{\dual Q}*{(x,y)}}
	\formula[y=0]{\exists*{z.~}P*{(f(z))}\*Q*{(z,t)}}
	\Vlink[red]{3,2}
	\Vlink[blue]{5,4}
\end{tikzpicture}
\]

Correctness and sequentialization for \all1 proof nets will be through \defn{coalescence}~\cite{Heijltjes-Hughes-2015,Hughes-Heijltjes-2016}, the additive counterpart to \defn{contractibility} for multiplicative linear logic~\cite{Danos-1990}. It is a simple graph rewrite relation that pushes links from the leaves towards the roots of a sequent, that is effectively top-down sequentialization (like its multiplicative counterpart~\cite{Guerrini-Masini-2001}). Figure~\ref{fig:main example} illustrates the process for the witness net above. In the example, a link between subformulas $A$ and $B$ carrying a substitution map $\sigma$ represents a sequent $\seq{A\sigma}{B\sigma}$. Each step (labelled $\star S$) corresponds to a proof rule (labelled $\star R$), by which the coalescence sequence generates the sequent proof.

In addition, we will present a geometric correctness condition for witness nets, and define composition as direct,

%In the former, to join two slices the witness assignment must be the same; in the latter, witness assignments must be compatible, so that a common, more general witness map exists.

\begin{figure}
\[
\begin{array}{c}
\begin{array}{@{}c@{~}c@{\qquad}c@{}}
  \vc{
  \begin{tikzpicture}[net]
	\formula[y=1.2]{\forall*{x.}\exists*{y.~}{\dual P}*{(y)}\+{\dual Q}*{(x,y)}}
	\formula[y=0]{\exists*{z.~}P*{(f(z))}\*Q*{(z,t)}}
	\Vlink[red,l,label={$\scriptstyle[f(x)/y\,,\,x/z]$~~}]{3,2}
	\Vlink[blue,r,label={~$\scriptstyle[t/y\,,\,x/z]$}]{5,4}
  \end{tikzpicture}}
&&
  \vc{
  \begin{tikzpicture}[net]
	\formula[y=1.2]{\forall*{x.}\exists*{y.~}{\dual P}*{(y)}\+{\dual Q}*{(x,y)}}
	\formula[y=0]{\exists*{z.~}P*{(f(z))}\*Q*{(z,t)}}
	\Vlink[violet,l,label={$\scriptstyle\varnothing$~~}]{2,1}
  \end{tikzpicture}}
\\ \\[-5pt] \downarrow\scriptstyle{\srr+1} && \uparrow\scriptstyle\srr! \\ \\[-5pt]
  \vc{
  \begin{tikzpicture}[net]
	\formula[y=1.2]{\forall*{x.}\exists*{y.~}{\dual P}*{(y)}\+{\dual Q}*{(x,y)}}
	\formula[y=0]{\exists*{z.~}P*{(f(z))}\*Q*{(z,t)}}
	\Vlink[red,l,label={$\scriptstyle[f(x)/y\,,\,x/z]$~~~}]{4,2}
	\Vlink[blue,r,label={~$\scriptstyle[t/y\,,\,x/z]$}]{5,4}
  \end{tikzpicture}}
&&
  \vc{
  \begin{tikzpicture}[net]
	\formula[y=1.2]{\forall*{x.}\exists*{y.~}{\dual P}*{(y)}\+{\dual Q}*{(x,y)}}
	\formula[y=0]{\exists*{z.~}P*{(f(z))}\*Q*{(z,t)}}
	\Vlink[violet,l,label={$\scriptstyle\varnothing$~~}]{1,1}
  \end{tikzpicture}}  
\\ \\[-5pt] \downarrow\scriptstyle\srr?,f(x) && \uparrow\scriptstyle\srr?,x \\ \\[-5pt]  
  \vc{
  \begin{tikzpicture}[net]
	\formula[y=1.2]{\forall*{x.}\exists*{y.~}{\dual P}*{(y)}\+{\dual Q}*{(x,y)}}
	\formula[y=0]{\exists*{z.~}P*{(f(z))}\*Q*{(z,t)}}
	\Vlink[red,l,label={$\scriptstyle[x/z]$~~}]{2,2}
	\Vlink[blue,r,label={~$\scriptstyle[t/y\,,\,x/z]$}]{5,4}
  \end{tikzpicture}} 
&&  
  \vc{
  \begin{tikzpicture}[net]
	\formula[y=1.2]{\forall*{x.}\exists*{y.~}{\dual P}*{(y)}\+{\dual Q}*{(x,y)}}
	\formula[y=0]{\exists*{z.~}P*{(f(z))}\*Q*{(z,t)}}
	\Vlink[violet,l,label={$\scriptstyle[x/z]$~~~}]{[2]2,3}
  \end{tikzpicture}}
\\ \\[-5pt] \downarrow\scriptstyle{\srr+2} && \uparrow\scriptstyle\srr* \\ \\[-5pt]
  \vc{
  \begin{tikzpicture}[net]
	\formula[y=1.2]{\forall*{x.}\exists*{y.~}{\dual P}*{(y)}\+{\dual Q}*{(x,y)}}
	\formula[y=0]{\exists*{z.~}P*{(f(z))}\*Q*{(z,t)}}
	\Vlink[red,l,label={$\scriptstyle[x/z]$~~}]{2,2}
	\Vlink[blue,r,label={~$\scriptstyle[t/y\,,\,x/z]$}]{4,4}
  \end{tikzpicture}}
& \stackrel{\srr?,t}\scoal & 
  \vc{
  \begin{tikzpicture}[net]
	\formula[y=1.2]{\forall*{x.}\exists*{y.~}{\dual P}*{(y)}\+{\dual Q}*{(x,y)}}
	\formula[y=0]{\exists*{z.~}P*{(f(z))}\*Q*{(z,t)}}
	\Vlink[red,l,label={$\scriptstyle[x/z]$~~}]{2,2}
	\Vlink[blue,r,label={~~~$\scriptstyle[x/z]$}]{[4]2,4}
  \end{tikzpicture}}
\end{array}
\\ \\[20pt]
	\color{violet}\infer[\color{black}\Qrr!]{\color{black}\Seq{\forall x.\exists y.\dual P(y)\+\dual Q(x,y)}{\exists z.P(f(z))\*Q(z,t)}}{
	 \color{violet}\infer[\color{black}\Qrr?,x]{\color{black}\Seq{\exists y.\dual P(y)\+\dual Q(x,y)}{\exists z.P(f(z))\*Q(z,t)}}{
	  \color{violet}\infer[\color{black}\Qrr*]{\color{black}\Seq{\exists y.\dual P(y)\+\dual Q(x,y)}{P(f(x))\*Q(x,t)}}{
	   \color{red}\infer[\color{black}\Qrr?{f(x)}]{\color{black}\Seq{\exists y.\dual P(y)\+\dual Q(x,y)}{P(f(x))}}{
	    \color{red}\infer[\color{black}\Qrr+1]{\color{black}\Seq{\dual P(f(x))\+\dual Q(x,f(x))}{P(f(x))}}{
	     \color{red}\infer[\color{black}\Qrr1]{\color{black}\Seq{\dual P(f(x))}{P(f(x))}}{}
	   }}&
	   \color{blue}\infer[\color{black}\Qrr?t]{\color{black}\Seq{\exists y.\dual P(y)\+\dual Q(x,y)}{Q(x,t)}}{
	    \color{blue}\infer[\color{black}\Qrr+2]{\color{black}\Seq{\dual P(t)\+\dual Q(x,t)}{Q(x,t)}}{
	     \color{blue}\infer[\color{black}\Qrr1]{\color{black}\Seq{\dual Q(x,t)}{Q(x,t)}}{}
	}}}}}
\end{array}
\]
\caption{A coalescence and sequentialization example}
\label{fig:example}
\end{figure}

@@

At the heart of a theory of proof nets is the question of proof identity, of when two proofs should be considered equivalent, and should thus map onto the same proof net. 
%
%It can be approached semantically, by asking whether two proofs are identified in every model; 
%
The introduction of quantifiers creates an interesting issue: if two proofs differ by an immaterial choice of existential witness, should they be equivalent? For example, to prove the sequent $\seq{\exists x.P(x)}{\exists y.\dual P(y)}$ both quantifiers must receive \emph{the same} witness, as in the following two proofs, but any witness will do.
\[
	\vc{\infer={\seq{\exists x.P(x)}{\exists y.\dual P(y)}}{\infer{\seq{P(s)}{\dual P(s)}}{}}}
	\quad
	\stackrel?\equiv
	\quad
	\vc{\infer={\seq{\exists x.P(x)}{\exists y.\dual P(y)}}{\infer{\seq{P(t)}{\dual P(t)}}{}}}
\]
The issue is more pronounced where quantifiers are \emph{vacuous}, $\exists x.A$ where $x$ is not free in $A$. The proofs below left can only be distinguished even syntactically because the $\exists R$-rule makes the instantiating witness explicit.
%
Below right is an interesting intermediate variant: the witness $s$ or $t$ can be observed without explicit annotation in the $\exists R$-rule, but the choice is equally immaterial to the content of the proof as when the quantifier were vacuous.
\[
	\vc{\infer[\!\!\scriptstyle{\qrr?s}]{\seq{\exists x.P}{\dual P}}{\infer{\seq P{\dual P}}{}}}
	\quad
	\stackrel?\equiv
	\quad	
	\vc{\infer[\!\!\scriptstyle{\qrr?t}]{\seq{\exists x.P}{\dual P}}{\infer{\seq P{\dual P}}{}}}\qquad
	\vc{
	 \infer[\!\!\scriptstyle{\qrr?s}]{\seq{\exists x.P\+Q(x)}{\dual P}}{
	  \infer[\!\!\scriptstyle{\qrr+1}]{\seq{P\+Q(s)}{\dual P}}{\infer{\seq P{\dual P}}{}}
	}}
	\quad
	\stackrel?\equiv
	\quad	
	\vc{
	 \infer[\!\!\scriptstyle{\qrr?t}]{\seq{\exists x.P\+Q(x)}{\dual P}}{
	  \infer[\!\!\scriptstyle{\qrr+1}]{\seq{P\+Q(t)}{\dual P}}{\infer{\seq P{\dual P}}{}}
	}}
\]
In this paper we will not attempt to decisively settle the question of proof identity. Rather, we will present two natural and coherent notions of proof net, that represent two extremes. One, \defn{witness nets}, takes the choice of existential witnesses to be essential proof content, and rejects all three equivalences above. The other, \defn{unification nets}, removes any redundancy by rejecting explicit witnessing information altogether; instead, a most general witness for every quantifier is reconstructed by unification.

\subsection{Expansion trees}

>>>>

\subjclass{F.4.1: Mathematical logic--Proof theory}
\keywords{Linear logic, First-order logic, Proof nets}

@@

Witness nets are canonical for first-order additive sequent calculus. Unification nets in addition factor out any inessential choice for existential witnesses. 
Both notions of proof net are defined through coalescence (the additive version of multiplicative contractibility), and for witness nets a geometric correctness criterion is provided. Both feature a global composition operation.
\end{abstract}

@@

Additive linear logic $(\all)$ is the logic of product $(\*)$ and sum $(\+)$. Proofs, over a sequent $A\vdash B$ or, here, $\seq{\dual A}B$, are built from the canonical morphisms: projections, injections, and diagonals. These are captured naturally in \emph{proof nets} \cite{Girard-1987,Hughes-vanGlabbeek-2005}, which consist of a set of \emph{links} between subformulas in $\dual A$ and in $B$, satisfying a simple geometric correctness condition.
\[
\begin{array}{c@{\qquad}c@{\qquad}c@{\qquad}c}
	\begin{tikzpicture}[net]
		\formula[y=1.2]{{\dual a}}
		\formula[y=0]{a\*a}
		\Vlink[red]{[-2]1,1}
		\Vlink[red]{[2]1,3}
	\end{tikzpicture}
&
	\begin{tikzpicture}[net]
		\formula[y=1.2]{{\dual a}}
		\formula[y=0]{a\+b}
		\Vlink[red]{1,1}
	\end{tikzpicture}
&
	\begin{tikzpicture}[net]
		\formula[y=1.2]{{\dual a}\+{\dual b}}
		\formula[y=0]{b\*a}
		\Vlink[red]{1,3}
		\Vlink[red]{3,1}
	\end{tikzpicture}
&
	\begin{tikzpicture}[net]
		\formula[y=1.2]{*{\phantom(}{\dual a}*{\phantom)}
		 {{\+}}*({\dual b}*{\phantom)}
		 {{\+}}*{\phantom(}{\dual c}*)}
		\formula[y=0]{*(a*{\phantom)}
		{{\*}}*{\phantom(}b*)
		{{\*}}*{\phantom(}c*{\phantom)}}
		\Vlink[red]{1,1}
		\Vlink[red]{3,3}
		\Vlink[red]{5,5}
	\end{tikzpicture}
\\
	\text{\small diagonal}
&	\text{\small injection}
&	\text{\small symmetry}
&	\text{\small associativity}
\end{array}
\]
%
The logic is combinatorially rich, yet well-behaved and tractable: proof search \cite{Galmiche-Marion-1995,Galmiche-2000} and proof net correctness \cite{Heijltjes-Hughes-2015} are linear in $|A|\times|B|$ (with $|A|$ the size of the syntax tree of $A$); proof nets remain canonical and equally tractable when extended with the two units \cite{Heijltjes-2011,Heijltjes-Hughes-2015}; and the first-order case is merely NP-complete \cite{Heijltjes-Hughes-2015}.

At the same time, the ubiquity of its main operations, product and sum, mean the ideas and results it garners are widely applicable. It serves as a microcosmos for important ideas and observations: it fostered the \emph{connections method} of proof search \cite{Galmiche-2000}; it demonstrated the \emph{Blass problem} \cite{Abramsky-2003} of game semantics, that sequential strategies do not in general have associative composition; and it is at the root of the study of fixed points in linear logic \cite{Santocanale-2002a}. The logic describes two-way communication~\cite{Cockett-Santocanale-2009}, and is a core part of \emph{session types}~\cite{Honda-1993}.

{\all} is of course part of {\mall} (multiplicative-additive linear logic), and its lessons are clearly visible in the second author's canonical proof nets \cite{Hughes-vanGlabbeek-2005} for that fragment, as well as the first and second author's locally canonical \emph{conflict nets}~\cite{Hughes-Heijltjes-2016}. But its proof nets also appear as the \emph{skew fibrations} in the second author's \emph{combinatorial proofs} for classical logic~\cite{Hughes-2006-proofs}, and in the third authors study of the \emph{medial rule} for classical logic~\cite{str:RTA07}. To prepare the ground for the extension of \emph{combinatorial proofs} to first-order classical logic is one motivation for the% current presentation.

We present proof nets for first-order additive linear logic (\all1). The main content of first-order proof, and our central challenge, is the \emph{witness assignment} to existential quantifiers. 
%
The standard approach to evaluate a formula $\exists x.A$ is by a substitution $A[t/x]$, which destroys the subformula $A$.%
%
\footnote{Given $A[t/x]$ and $t$, it is not possible to unambiguously recover $A$, which may contain other occurrences of $t$ not originating in the substitution $[t/x]$.}
%
But the current, successful formulation of additive proof nets, as a set of links over a two-formula sequent, requires subformulas to remain intact.
%
Moreover, the same quantification $\exists x.A$ may occur in different \emph{slices} (branches of a proof above a product), with different witnesses, as below. But the \emph{formulas + links} design has only one instance of $\exists x.A$.
\[
\vc{
	\infer[\Qrr*]{\Seq{\exists x.A}{B\*C}}{
	 \infer[\Qrr?s]{\Seq{\exists x.A}B}{\Seq{A[s/x]}B}
	 &
	 \infer[\Qrr?t]{\Seq{\exists x.A}C}{\Seq{A[t/x]}C}
   }}
~.
\]
%
We advance two solutions. \defn{Witness nets} replace the \emph{implicit} substitutions of sequent proofs with \emph{explicit} substitutions, recorded at each link. \defn{Unification nets} omit existential witnesses altogether, and reconstruct them by unification. Respective examples are below.
\[
\begin{tikzpicture}[net]
	\formula[y=1.2]{\forall*{x.}\exists*{y.~}{\dual P}*{(y)}\+{\dual Q}*{(x,y)}}
	\formula[y=0]{\exists*{z.~}P*{(f(z))}\*Q*{(z,t)}}
	\Vlink[red,l,label={$\scriptstyle[f(x)/y\,,\,x/z]$~~}]{3,2}
	\Vlink[blue,r,label={~$\scriptstyle[t/y\,,\,x/z]$}]{5,4}
\end{tikzpicture}
\qquad\qquad
\begin{tikzpicture}[net]
	\formula[y=1.2]{\forall*{x.}\exists*{y.~}{\dual P}*{(y)}\+{\dual Q}*{(x,y)}}
	\formula[y=0]{\exists*{z.~}P*{(f(z))}\*Q*{(z,t)}}
	\Vlink[red]{3,2}
	\Vlink[blue]{5,4}
\end{tikzpicture}
\]
%
Correctness and sequentialization for \all1 proof nets will be through \defn{coalescence}~\cite{Heijltjes-Hughes-2015,Hughes-Heijltjes-2016}, the additive counterpart to \defn{contractibility} for multiplicative linear logic~\cite{Danos-1990}. It is a simple graph rewrite relation that pushes links from the leaves towards the roots of a sequent, that is effectively top-down sequentialization (like its multiplicative counterpart~\cite{Guerrini-Masini-2001}). Figure~\ref{fig:main example} illustrates the process for the witness net above. In the example, a link between subformulas $A$ and $B$ carrying a substitution map $\sigma$ represents a sequent $\seq{A\sigma}{B\sigma}$. Each coalescence step ($\star S$) corresponds to a proof rule ($\star R$), by which the sequence generates the sequent proof.

In unification nets, existential witnesses are reconstructed during coalescence. The initial substitution map assigned to a link on atomic formulas $a=P(s_1,\dots,s_n)$ and $b=\dual P(t_1,\dots,t_n)$ is the most general unifier (\mgu) of $a$ and $\dual b$ (the first step in Figure~\ref{fig:unification example}). Coalescence then proceeds as for witness nets, except when joining two links from different slices with a product step, ($\urr*$) in Figure~\ref{fig:unification example}. Here, witness nets require both links to carry the same substitution map, while for unification nets a common, more general map is generated by unification.

We include the following results for our \all1 proof nets. Witness nets are \emph{canonical} for the permutations of the sequent calculus, and we provide a geometric correctness condition. \emph{Composition} of unification nets (over $\seq AB$ and $\seq{\dual BC}$ into one over $\seq AC$) is remarkedly simple, as the relational composition of both sets of links. For composition of witness nets, in addition the witness assignments of links must be composed, which is done through a simple process of \emph{interaction + hiding} similar to those of game semantics \cite{Mimram-2011}.

A companion paper \cite{Hughes-2018} explores unification nets for first-order \emph{multiplicative} linear logic, reporting drastic complexity improvements and equally straightforward proof net composition.

% Hyland-Ong-2000,Abramsky-Jagadeesan-Malacaria-2000

\begin{figure}
\[
\begin{array}{c}
\begin{array}{@{}c@{~}c@{\qquad}c@{}}
  \vc{
  \begin{tikzpicture}[net]
	\formula[y=1.2]{\forall*{x.}\exists*{y.~}{\dual P}*{(y)}\+{\dual Q}*{(x,y)}}
	\formula[y=0]{\exists*{z.~}P*{(f(z))}\*Q*{(z,t)}}
	\Vlink[red,l,label={$\scriptstyle[f(x)/y\,,\,x/z]$~~}]{3,2}
	\Vlink[blue,r,label={~$\scriptstyle[t/y\,,\,x/z]$}]{5,4}
  \end{tikzpicture}}
&&
  \vc{
  \begin{tikzpicture}[net]
	\formula[y=1.2]{\forall*{x.}\exists*{y.~}{\dual P}*{(y)}\+{\dual Q}*{(x,y)}}
	\formula[y=0]{\exists*{z.~}P*{(f(z))}\*Q*{(z,t)}}
	\Vlink[violet,l,label={$\scriptstyle\varnothing$~~}]{1,1}
  \end{tikzpicture}} 
\\ \\[-5pt] \downarrow\scriptstyle{\srr+1} && \uparrow\scriptstyle\srr! \\ \\[-5pt]
  \vc{
  \begin{tikzpicture}[net]
	\formula[y=1.2]{\forall*{x.}\exists*{y.~}{\dual P}*{(y)}\+{\dual Q}*{(x,y)}}
	\formula[y=0]{\exists*{z.~}P*{(f(z))}\*Q*{(z,t)}}
	\Vlink[red,l,label={$\scriptstyle[f(x)/y\,,\,x/z]$~~~}]{4,2}
	\Vlink[blue,r,label={~$\scriptstyle[t/y\,,\,x/z]$}]{5,4}
  \end{tikzpicture}}
&&
  \vc{
  \begin{tikzpicture}[net]
	\formula[y=1.2]{\forall*{x.}\exists*{y.~}{\dual P}*{(y)}\+{\dual Q}*{(x,y)}}
	\formula[y=0]{\exists*{z.~}P*{(f(z))}\*Q*{(z,t)}}
	\Vlink[violet,l,label={$\scriptstyle\varnothing$~~}]{2,1}
  \end{tikzpicture}} 
\\ \\[-5pt] \downarrow\scriptstyle\srr?,f(x) && \uparrow\scriptstyle\srr?,x \\ \\[-5pt]  
  \vc{
  \begin{tikzpicture}[net]
	\formula[y=1.2]{\forall*{x.}\exists*{y.~}{\dual P}*{(y)}\+{\dual Q}*{(x,y)}}
	\formula[y=0]{\exists*{z.~}P*{(f(z))}\*Q*{(z,t)}}
	\Vlink[red,l,label={$\scriptstyle[x/z]$~~}]{2,2}
	\Vlink[blue,r,label={~$\scriptstyle[t/y\,,\,x/z]$}]{5,4}
  \end{tikzpicture}} 
&&  
  \vc{
  \begin{tikzpicture}[net]
	\formula[y=1.2]{\forall*{x.}\exists*{y.~}{\dual P}*{(y)}\+{\dual Q}*{(x,y)}}
	\formula[y=0]{\exists*{z.~}P*{(f(z))}\*Q*{(z,t)}}
	\Vlink[violet,l,label={$\scriptstyle[x/z]$~~~}]{[2]2,3}
  \end{tikzpicture}}
\\ \\[-5pt] \downarrow\scriptstyle{\srr+2} && \uparrow\scriptstyle\srr* \\ \\[-5pt]
  \vc{
  \begin{tikzpicture}[net]
	\formula[y=1.2]{\forall*{x.}\exists*{y.~}{\dual P}*{(y)}\+{\dual Q}*{(x,y)}}
	\formula[y=0]{\exists*{z.~}P*{(f(z))}\*Q*{(z,t)}}
	\Vlink[red,l,label={$\scriptstyle[x/z]$~~}]{2,2}
	\Vlink[blue,r,label={~$\scriptstyle[t/y\,,\,x/z]$}]{4,4}
  \end{tikzpicture}}
& \stackrel{\srr?,t}\scoal & 
  \vc{
  \begin{tikzpicture}[net]
	\formula[y=1.2]{\forall*{x.}\exists*{y.~}{\dual P}*{(y)}\+{\dual Q}*{(x,y)}}
	\formula[y=0]{\exists*{z.~}P*{(f(z))}\*Q*{(z,t)}}
	\Vlink[red,l,label={$\scriptstyle[x/z]$~~}]{2,2}
	\Vlink[blue,r,label={~~~$\scriptstyle[x/z]$}]{[4]2,4}
  \end{tikzpicture}}
\end{array}
\\ \\[20pt]
	\color{violet}\infer[\color{black}\Qrr!]{\color{black}\Seq{\forall x.\exists y.\dual P(y)\+\dual Q(x,y)}{\exists z.P(f(z))\*Q(z,t)}}{
	 \color{violet}\infer[\color{black}\Qrr?,x]{\color{black}\Seq{\exists y.\dual P(y)\+\dual Q(x,y)}{\exists z.P(f(z))\*Q(z,t)}}{
	  \color{violet}\infer[\color{black}\Qrr*]{\color{black}\Seq{\exists y.\dual P(y)\+\dual Q(x,y)}{P(f(x))\*Q(x,t)}}{
	   \color{red}\infer[\color{black}\Qrr?{f(x)}]{\color{black}\Seq{\exists y.\dual P(y)\+\dual Q(x,y)}{P(f(x))}}{
	    \color{red}\infer[\color{black}\Qrr+1]{\color{black}\Seq{\dual P(f(x))\+\dual Q(x,f(x))}{P(f(x))}}{
	     \color{red}\infer[\color{black}\Qrr1]{\color{black}\Seq{\dual P(f(x))}{P(f(x))}}{}
	   }}&
	   \color{blue}\infer[\color{black}\Qrr?t]{\color{black}\Seq{\exists y.\dual P(y)\+\dual Q(x,y)}{Q(x,t)}}{
	    \color{blue}\infer[\color{black}\Qrr+2]{\color{black}\Seq{\dual P(t)\+\dual Q(x,t)}{Q(x,t)}}{
	     \color{blue}\infer[\color{black}\Qrr1]{\color{black}\Seq{\dual Q(x,t)}{Q(x,t)}}{}
	}}}}}
\end{array}
\]
\caption{A coalescence and sequentialization example}
\label{fig:main example}
\end{figure}

\begin{figure}
\[
\vc{\begin{tikzpicture}[net]
	\formula[y=1.2]{{\dual P}*{(x,x)}}
	\formula[y=0]{P*{(y,s)}\*P*{(t,z)}}
	\Vlink[red]{[-2]1,1}
	\Vlink[blue]{[2]1,3}
\end{tikzpicture}}
\quad\ucoal\quad
\vc{\begin{tikzpicture}[net]
	\formula[y=1.2]{{\dual P}*{(x,x)}}
	\formula[y=0]{P*{(y,s)}\*P*{(t,z)}}
	\Vlink[red,l,label={$\scriptstyle[s/y,s/x]$~~}]{[-2]1,1}
	\Vlink[blue,r,label={~~$\scriptstyle[t/z,t/x]$}]{[2]1,3}
\end{tikzpicture}}
\quad\stackrel{\urr*}\ucoal\quad
\vc{\begin{tikzpicture}[net]
	\formula[y=1.2]{{\dual P}*{(x,x)}}
	\formula[y=0]{P*{(y,s)}\*P*{(t,z)}}
	\Vlink[violet,r,label={~~$\scriptstyle[u/z,u/y,u/x]$}]{1,2}
\end{tikzpicture}}
\]
\caption{Coalescence with unification; $u$ is the smallest term unifying $s$ and $t$} 
\label{fig:unification example}
\end{figure}

@@

At the heart of a theory of proof nets is the question of \emph{proof identity}: when are two proofs equivalent? The answer determines which proofs should map onto the same proof net. The introduction of quantifiers creates an interesting issue: if two proofs differ by an immaterial choice of existential witness, should they be equivalent? For example, to prove the sequent $\seq{\exists x.P(x)}{\exists y.\dual P(y)}$ both quantifiers must receive \emph{the same} witness, as in the following two proofs, but any witness will do.
\[
	\vc{\infer={\seq{\exists x.P(x)}{\exists y.\dual P(y)}}{\infer{\seq{P(s)}{\dual P(s)}}{}}}
	\quad
	\stackrel?\equiv
	\quad
	\vc{\infer={\seq{\exists x.P(x)}{\exists y.\dual P(y)}}{\infer{\seq{P(t)}{\dual P(t)}}{}}}
\]
The issue is more pronounced where quantifiers are \emph{vacuous}, $\exists x.A$ where $x$ is not free in $A$. The proofs below left can only be distinguished even syntactically because the $\exists R$-rule makes the instantiating witness explicit. Below right is an interesting intermediate variant: the witness $s$ or $t$ can be observed without explicit annotation in the $\exists R$-rule, but the choice is equally immaterial to the content of the proof as when the quantifier were vacuous.
\[
	\vc{\infer[\!\!\scriptstyle{\qrr?s}]{\seq{\exists x.P}{\dual P}}{\infer{\seq P{\dual P}}{}}}
	\quad
	\stackrel?\equiv
	\quad	
	\vc{\infer[\!\!\scriptstyle{\qrr?t}]{\seq{\exists x.P}{\dual P}}{\infer{\seq P{\dual P}}{}}}\qquad
	\vc{
	 \infer[\!\!\scriptstyle{\qrr?s}]{\seq{\exists x.P\+Q(x)}{\dual P}}{
	  \infer[\!\!\scriptstyle{\qrr+1}]{\seq{P\+Q(s)}{\dual P}}{\infer{\seq P{\dual P}}{}}
	}}
	\quad
	\stackrel?\equiv
	\quad	
	\vc{
	 \infer[\!\!\scriptstyle{\qrr?t}]{\seq{\exists x.P\+Q(x)}{\dual P}}{
	  \infer[\!\!\scriptstyle{\qrr+1}]{\seq{P\+Q(t)}{\dual P}}{\infer{\seq P{\dual P}}{}}
	}}
\]
In this paper we will not attempt to settle the question of proof identity. Rather, our two notions of proof net each represent a natural and coherent perspective, at either end of the spectrum. \emph{Witness nets} make all existential witnesses explicit, including those to vacuous quantifiers, rejecting all three equivalences above. \emph{Unification nets} leave all witnesses implicit, thus identifying all proofs modulo witness assignment, and validating all three equivalences.

\subsection{Monomial nets and Expansion Trees}

Proof nets with additives and quantifiers exist as \emph{monomial nets} \cite{Girard-1996}, where slices are managed explicitly with \emph{monomial weights}. These nets are not generally canonical: they admit the permutation (and duplication) of proof rules past implicit \emph{contractions}. We believe (though have not technically verified) that additive monomial nets include, and could be restricted to, canonical forms. These would be equivalent to the following approach.
